% Copyright (c) 2016 The ALF project.
% This is a part of the ALF project documentation.
% The ALF project documentation by the ALF contributors is licensed
% under a Creative Commons Attribution-ShareAlike 4.0 International License.
% For the licensing details of the documentation see license.CCBYSA.

% !TEX root = doc.tex

%-------------------------------------------------------------------------------------
\section*{Acknowledgments} 
%-------------------------------------------------------------------------------------

%\red{To be updated}
%
%Future: - Improved overall usability
%- Improved I.O., HDF5
%- Improved post-processing
%- Implementation with classes
%- Trotter options
%- Time-dependent Hamiltonians\\
%- Hybrid Monte Carlo with exact forces
%- Rényi entropies
%- Interaction expansion (CT-INT).

We are very grateful to  B.~Danu, S.~Beyl, M.~Hohenadler,  M.~Raczkowski, T.~Sato,  M.~Ulybyshev, Z.~Wang, and M.~Weber for their constant support during the development of this project.
We equally thank G.~Hager, M.~Wittmann, and G.~Wellein for useful discussions and overall support.
FFA would also like to thank T.~Lang  and Z.~Y.~Meng for developments of the auxiliary field code as well as to T.~Grover. 
MB, FFA  and FG thank the Bavarian Competence Network for Technical and Scientific High Performance Computing (KONWIHR) for financial support.
FG, JH, and JS thank the SFB-1170 for  financial support under projects Z03 and C01.
F.P.T is funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) -- project number 414456783.
JSEP  thanks the DFG for financial support under the project AS120/14-1, dedicated to the further development of the ALF library.
Part of the optimization of the code was carried out during  the  Porting and Tuning Workshop 2016 offered by the Forschungszentrum J\"ulich.
Calculations performed to extensively test this package were carried out both on  SuperMUC-NG at the  Leibniz Supercomputing Centre and on  JURECA  \cite{Jureca16} at the J\"ulich Supercomputing Centre.  We thank both institutions for the generous allocation of computing time.
 %The authors gratefully acknowledge the computing time granted by the John von Neumann Institute for Computing (NIC) and provided on the supercomputer JURECA \cite{Jureca16} at Jülich Supercomputing Centre (JSC). The authors gratefully acknowledge the Gauss Centre for Supercomputing e.V. (www.gauss-centre.eu) for funding this project by providing computing time on the GCS Supercomputer SuperMUC at the Leibniz Supercomputing Centre (LRZ, www.lrz.de).
